\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\newtheorem{theorem}{Theorem}							% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{First Logic set}
\author{The Author}
%\date{}							% Activate to display a given date or no date
\newtheorem{axiom}{Axiom}
\newtheorem{lemma}{Lemma}
\newtheorem{puzzle}{Puzzle}

\newtheorem{definition}{Definition}
\begin{document}
\maketitle
%\section{}
%\subsection{}

\begin{lemma}
Sow that the following are equivalent (here $P Q R$ are propositions)
\begin{enumerate}	
\item $(P \rightarrow Q) \land (P \rightarrow R)$ and $P \rightarrow (Q\land R)$
\item $(P \rightarrow Q) \lor (P \rightarrow R)$ and $P \rightarrow (Q\lor R)$

\item $R\rightarrow(Q\rightarrow R)$ and $(P\land Q) \rightarrow R$.
\item $(P \rightarrow R) \land (Q \rightarrow R)$ and $(P \lor Q) \rightarrow R$
\item $(P \rightarrow R) \lor (Q \rightarrow R)$ and $(P \land Q) \rightarrow R$
\end{enumerate}


\end{lemma}

\begin{lemma}

\begin{enumerate}
\item $(P \lor Q) \land \neg  P \rightarrow Q$
\item $\neg Q ? (P \rightarrow Q) \rightarrow  \neg P$
\item $ C \land (A \rightarrow B) \land (C \rightarrow ( A \rightarrow \neg B)) \rightarrow \neg A$
\item $ (P \lor  Q) \land (\neg P \lor R) \rightarrow Q \lor R$.
\end{enumerate}

\end{lemma}



\begin{lemma} Prove or disprove
\begin{enumerate}
\item Either sales or expenses will go up. If sales go up then the boss will be happy. 
If expenses go up then the boss will be unhappy. Therefore sales and expenses will not both go up.

\item If the tax rate and unemployment rate go up, then there will be a recession. 
If the GDP goes up then there will not be a recession, The GDP and taxes are both going up. 
Therefore the unemployment is not going up.

\item The warning light will come on if and only if the pressure is high and the relief valve is clogged. 
The relief valve is not clogged. therefore, the warning light will come on if and only if the pressure is too high.
\end{enumerate}
\end{lemma}

\begin{puzzle} What can you deduce from the following statements?
\begin{enumerate}
\item All babies are illogical.
\item Nobody is despised who can manage a crocodile.
\item Illogical persons are despised.
\end{enumerate}
\end{puzzle}


\begin{puzzle} What can you deduce from the following statements?
\begin{enumerate}
\item No ducks waltz.
\item No officers ever decline to waltz.
\item All my poultry are ducks.
\end{enumerate}
\end{puzzle}



\begin{puzzle} What can you deduce from the following statements?
\begin{enumerate}
\item Things sold in the street are of no great value.
\item Nothing but rubbish can be had for a song.
\item Eggs of the Great Auk are very valuable.
\item It is only what is sold in the street that is really rubbish.

\end{enumerate}
\end{puzzle}
\begin{puzzle} What can you deduce from the following statements?
\begin{enumerate}
\item All writers, who understand human nature, are clever.
\item No one is a true poet unless he can stir the hearts of men.
\item Shakespeare wrote ?Hamlet?.
\item No writer, who does not understand human nature, can stir the hearts of men.
\item None but a true poet could have written ?Hamlet?.

\end{enumerate}
\end{puzzle}


\end{document}  